Have you forgotten?' Conversion experiences in lists

### 2016-10-21 lists_conversion_experience.rmd
fixing up formatting and adding new figures

```{r setup, cache=FALSE, include=FALSE}
library(knitr)
library(tint)
knitr::opts_chunk$set(tidy = FALSE, cache.extra = packageVersion('tint'))
options(htmltools.dir.version = FALSE)
output <- opts_knit$get("rmarkdown.pandoc.to")
opts_chunk$set(warning=FALSE)
opts_chunk$set(message=FALSE)
opts_chunk$set(error=TRUE)
```

@@

Amidst the predictive conditions of contemporary media-culture, the inevitable incompleteness of lists is both a problem and a potential. 
This paper investigates practices of list-making associated with online grocery shopping at a  UK and global supermarket chain, Tesco. 
It will analyse how a recommendation system suggests a small number of grocery items of personal relevance to each of the millions of online grocery shoppers. 
The predictive modelling practices aim to up-sell and cross-sell some of the 250,000 items in the supermarket's inventory. 
I find it interesting to approach the list-making practices  of the grocery recommendation system as a mundane case of the predictive intermediaries operating in many places. 
The transformation or conversion of listing into predicting can be analysed in recommendation systems that extend existing lists with potential new items, thus transforming the temporal and agential fabric of the list.
While all written lists mix past and future actions, actual and potential events, their predictive re-writing in recommendation systems adds a new operational dynamic.
In this operational dynamic, the incompleteness of lists, their propensity to expand or change, becomes more important than their capacity to exhaustively enumerate things. 
The conversion of lists from finite to in-definite in recommendation systems suggests we might need to re-conceptualise how we think about social order and structure in media cultures.

@@

I saw recently a presentation about online grocery shopping at Tesco, a large supermarket chain based in the UK. The lists in this setting are recommendations for further grocery purchases under the title of 'Have you forgotten?'. When you shop for groceries online, a list of 'Have you forgotten' recommendations appear  at the checkout stage. While the title 'have you forgotten' implies that you will be shown items that could have been or are usually on a shopping list, the working assumption is a temporal one -- the list is incomplete either because you have forgotten to list something, or you didn't know what should be on your list, perhaps because you have never even thought of putting that thing on your list.

@@

[^12]: The problem of incomplete lists reminds me of the game 'I went shopping and I bought ... ' that adults sometimes play with children.  The game is about remembering a well-ordered list of items - 'a' is apple, 'b' is for banana, 'c' is for carrot.  The point of the game is to use the well-ordered set of labels provided by the sequence of letters in the alphabet to remember a random and perhaps very diverse list of items. The longer the game continues, the harder it is to remember since the shopping list grows longer. At the same time, each player repeats the list, and this  helps memorisation. One trick in playing this game is to choose items that are quite unexpected or increasingly random. Rather than `ham`, list `halva`; rather than `fish`, add `face cream.` This is especially useful for adults in beating children. The children have better memories for items they know, but adults know more unexpected items. On the other hand, very unexpected items are quite easy to remember and tend to even the score in the favour of the children. Balancing between the improbable and memorable, between a predictable list like 'apples, bananas, carrots' and an unusual list such as 'agave syrup, boccadilla, cilantro' is complicated. Something similar applies to recommendation systems. The system can remember very many things. It may know everything you bought in the last few years on that site, and it can easily see what you would normally buy and add those as suggestions.

@@

To this end, I will be seeking to simulate the workings of the recommendation system by generating queries, and exploring results from  a very large supermarket chain database.

```{r simulate_basket, echo=FALSE, cache=TRUE, fig.margin=TRUE}
library(dplyr)
library(ggplot2)
library(readr)

@@

ggplot(df, aes(x=price,y=ContentsQuantity)) + geom_point() + theme(axis.text=element_text(size=10)) 
df[df$price>=20,'name']
```

@@

How would I characterise this shift? Use it to emphasize why attention to the algorithm is not enough -- it is the model that counts
A little history of recommender systems -- uses Varian 1997 for a start. But has no real reference list. Not sure  
Also can use Schafer -- back to Association Rule Mining from early 1990s

@@

@@

```{r arules, echo=TRUE, fig.margin=TRUE}
library(arules)
library(arulesViz)
library(datasets)
data(Groceries)
itemFrequencyPlot(Groceries,topN=20,type="absolute")
# Get the rules
rules <- apriori(Groceries, parameter = list(supp = 0.001, conf = 0.8))

# Show the top 5 rules, but only 2 digits
options(digits=2)
inspect(rules[1:5])
summary(rules)

rules<-sort(rules, by="confidence", decreasing=TRUE)
inspect(rules[1:5])
rules<-apriori(data=Groceries, parameter=list(supp=0.001,conf = 0.15,minlen=2), 
           appearance = list(default="rhs",lhs="whole milk"),
           control = list(verbose=F))
rules<-sort(rules, decreasing=TRUE,by="confidence")
inspect(rules[1:5])  
items = Groceries@itemInfo
```

The association rule-based model for the `Groceries` dataset suggests some of the characteristic ordering practices of data-mining approaches to lists. 
The association between items is expressed in the form of _rules_. These rules, a term derived from decision support literature, define sets of associations between items. As a set theory-based practice, the association rules define likely sets of items or shopping baskets in terms of decreasing probability.   `r inspect(rules[1])` is a stronger association than `r inspect(rules[2])`. So any recommendation in the 'Have You Forgotten' list should probably put them in that order.

Association rules and the `apriori` algorithm present social order in terms of sets of common associations. This means that that the algorithm can be understood in terms of the set-based multiples proposed by Alain Badiou:

@@

Patel observed that the 'Have you forgotten' lists matter to people, and we can see why in the light of this proliferation of things in the list. A `r total` list of items is quite hard to list let alone read. 
Patel said that some people log on and go straight to the checkout, allowing the recommendation system to supply them with the suggestions that they then move to the shopping basket.
This is the online equivalent of using an old shopping list and only adding new items when needed. 
It is also a strategy to avoid having to view too many lists, since the only way Tesco can show people what they might buy is by showing them lists of things (e.g. chicken products), but these lists are often too long to work with. For instance, Tesco has more than 1000 chicken products on its site.   
People inhabit lists and work with them no matter how they are made; thus vindicating ethnomethodologists assertion that social order is the product of ongoing sense-making practices. 
But it also suggests a transformation in the way that lists exist in the world. They are less the product of a discrete agent than an effect of an environment teeming with possible combinations and complementaries, a surplus of possible purchases that the supermarket needs to continually maintain but also filter in order to make it possible for something to happen.

@@

Patel described the data they on grocery purchases. 'We have 52 weeks of purchase data for each customer.' 
We know that the construction of lists has been transformed by recommendation systems.
These transformations have been widely discussed in the context of big data, where questions of who has how much data are usually most prominent.
In terms of the ways that lists order the world, the amount of data is not so amount as the way in which things in the world take on a list-like existence. 
I want to dwell on this idea for a while. We have seen two different ways in which lists might be constructed in a recommender system. 
The first uses history and existence of previous lists, including the customer's own list, to populate a new list. 
The second uses many shopping lists to identify associations between things and then uses the strongest associations to recommend additions to a current list. 
The 'relevancy model,' however, tries to do something different. 
The problem that the proliferation of things on the shelves of supermarket or grocery warehouse confronts is partly combinatorial.
With  `r total` products (actually Patel mentions 200,000 products) can be combined in truly astronomical numbers of combinations.
If a typical shopping list has 20 items, then there are `r format(choose(total, 20), 2)` possible combinations.
Some of this vast number of possible shopping lists is highly unlikely in terms of shopping lists and histories of shopping, but potentially highly interesting and value as recommendation.

@@

The fact that so little was said about the models themselves suggests that they are now largely taken for granted in recommendation systems. The models she mentioned are all classifiers, predictive models that treat a set of variables in a datasets as a vector space in which good classification means finding a way of defining a decision boundary between points that belong to different classes of things. Here the classes of things are binary: `recommended` or `not recommended.` Some classifier models, including logistic regression, generate a probability rather than simply a classification.

@@

Predictive models, we might say, encounter their own surplus of sensibility in the data as they search for the most accurate predictions. A  very brief comment from Patel -- 'we have lots of zeros' -- points to the problem here. Each of the millions of Tesco online grocery shoppers has bought some selection of the 200,000 products. But the product-customer matrix, the basic vector-space in which all recommendation systems operate, remain very sparse and unpopulated.

@@

The discussion of imperfection was brief - it incorporated the notion of recommendations weighted by price and the notion of precision (or the proportion of products actually purchased.). Patel simply said that this metric was imperfect, but without elaborating why. Instead she moved to describe some of the ways in which the recommendation system worked in practice.

@@

Random allocation of customers to the four categories (Test A, Test B, Control A, Control B) adds another layer of probability to the recommendation system. This is done in the name of statistical validation of the effects or the uplift of the model. It also suggests that the effects of the model cannot be known in advance -- they can only be observed experimentally. Does this experimentality of the list point more to the propensities that Hansen says saturate experience?

It is perhaps significant that this random allocation of customers to control and test groups occurs without any relation to the particular profile or propensities of the customer. It seeks to statistically validate the effects -- the *uplift* -- of the model on  conversion events by allowing the effects of the model on what people do to be measured. The lift or uplift refers to a change in events associated with the same groups of people. This effectively is an experiment in inhabiting two different worlds.  This sets up a feedback look between the predictive model, the logistic regression, and the operation of the recommendation system.[^122] Once the experimental uplift methodology is in place, the recommendation system can constantly modify the recommendation list in light of subsequent data on the uplift. Predictions, then, are not bound by fixed parameters of the model.

@@

Hansen refers to an ontological transformation in which predictive media or recommendation systems would be seen as the closed realisation of the actually open propensities of the world. By understanding the world in terms of propensities, he argues, we have a change to engage with contemporary predictive media.

>>>>

```{r setup, cache=FALSE, include=FALSE}
library(knitr)
library(tint)
knitr::opts_chunk$set(tidy = FALSE, cache.extra = packageVersion('tint'))
options(htmltools.dir.version = FALSE)
output <- opts_knit$get("rmarkdown.pandoc.to")
opts_chunk$set(warning=FALSE)
opts_chunk$set(message=FALSE)
opts_chunk$set(error=TRUE)
opts_chunk$set(cache=TRUE)
```

@@

Amidst the predictive conditions of contemporary media-culture, the inevitable incompleteness of lists is both a problem and a potential. 
This paper investigates practices of list-making associated with online grocery shopping at a  UK and global supermarket chain, Tesco. 
It will analyse how a recommender system suggests a small number of grocery items of personal relevance to each of the millions of online grocery shoppers. 
The predictive modelling practices aim to up-sell and cross-sell some of the 250,000 items in the supermarket's inventory. 
I find it interesting to approach the list-making practices  of the grocery recommender system as a mundane case of the predictive intermediaries operating in many places. 
The transformation or conversion of listing into predicting can be analysed in recommender systems that extend existing lists with potential new items, thus transforming the temporal and agential fabric of the list.
While all written lists mix past and future actions, actual and potential events, their predictive re-writing in recommender systems adds a new operational dynamic.
In this operational dynamic, the incompleteness of lists, their propensity to expand or change, becomes more important than their capacity to exhaustively enumerate things. 
The conversion of lists from finite to in-definite in recommender systems suggests we might need to re-conceptualise how we think about social order and structure in media cultures.

@@

I saw recently a presentation about online grocery shopping at Tesco, a large supermarket chain based in the UK. The lists in this setting are recommendations for further grocery purchases under the title of 'Have you forgotten?'. When you shop for groceries online, a list of 'Have you forgotten' recommendations appear  at the checkout stage. While the title 'have you forgotten' implies that you will be shown items that could have been or are usually on a shopping list, the working assumption is a temporal one -- the list is incomplete either because you have forgotten to list something, or you didn't know what should be on your list, perhaps because you have never even thought of putting that thing on your list.
- making some changes, including a slide of early recommendation systems from 1990s such as described in ~/archive/data_intensive/machine_learning/schafer_recommenders_ecommerce_2001.pdf
-

@@

[^12]: The problem of incomplete lists reminds me of the game 'I went shopping and I bought ... ' that adults sometimes play with children.  The game is about remembering a well-ordered list of items - 'a' is apple, 'b' is for banana, 'c' is for carrot.  The point of the game is to use the well-ordered set of labels provided by the sequence of letters in the alphabet to remember a random and perhaps very diverse list of items. The longer the game continues, the harder it is to remember since the shopping list grows longer. At the same time, each player repeats the list, and this  helps memorisation. One trick in playing this game is to choose items that are quite unexpected or increasingly random. Rather than `ham`, list `halva`; rather than `fish`, add `face cream.` This is especially useful for adults in beating children. The children have better memories for items they know, but adults know more unexpected items. On the other hand, very unexpected items are quite easy to remember and tend to even the score in the favour of the children. Balancing between the improbable and memorable, between a predictable list like 'apples, bananas, carrots' and an unusual list such as 'agave syrup, boccadilla, cilantro' is complicated. Something similar applies to recommender systems. The system can remember very many things. It may know everything you bought in the last few years on that site, and it can easily see what you would normally buy and add those as suggestions.

@@

To this end, I will be seeking to simulate the workings of the recommender system by generating queries, and exploring results from  a very large supermarket chain database.

```{r simulate_basket, echo=FALSE, cache=TRUE, fig.margin=TRUE, fig.cap='Prices and quantities of things in the world'}
library(dplyr)
library(ggplot2)
library(readr)
library(knitr)

@@

ggplot(df, aes(x=price,y=ContentsQuantity)) + geom_point() + theme(axis.text=element_text(size=10)) 
kable(df[df$price>=20,'name'])
```

@@

How would I characterise this shift? Use it to emphasize why attention to the algorithm is not enough -- it is the model that counts
A little history of recommender systems -- uses Varian 1997 for a start. But has no real reference list. 
Also can use Schafer -- back to Association Rule Mining from early 1990s

@@

![Early e-commerce recommender system architecture](figure/schafer_recommendation_2001.png)

![The advent of recommender systems](figure/resnick_varian_page1_197.png)

@@

```{r arules, cache=TRUE, echo=TRUE, fig.margin=TRUE}
library(knitr)
library(arules)
library(arulesViz)
library(datasets)
data(Groceries)
itemFrequencyPlot(Groceries,topN=20,type="absolute")
rules <- apriori(Groceries, parameter = list(supp = 0.001, conf = 0.8))

options(digits=2)
rules<-sort(rules, by="confidence", decreasing=TRUE)
rules<-apriori(data=Groceries, parameter=list(supp=0.001,conf = 0.15,minlen=2), 
           appearance = list(default="rhs",lhs="whole milk"),
           control = list(verbose=F))
rules<-sort(rules, decreasing=TRUE,by="confidence")
kable(inspect(rules[1:5]))
items = Groceries@itemInfo
```

The association rule-based model for the `Groceries` dataset suggests some of the characteristic ordering practices of 1990s-style data-mining approaches to lists. 
The association between items is expressed in the form of _rules_. These rules, a term derived from decision support literature, define sets of associations between items. As a set theory-based practice, the association rules define likely sets of items or shopping baskets in terms of decreasing probability.   `r inspect(rules[1])` is a stronger association than `r inspect(rules[2])`. So any recommendation in the 'Have You Forgotten' list should probably put them in that order.

Association rules and the `apriori` algorithm present social order in terms of sets of common associations that are already weighted by frequency and hence probability. This means that that the algorithm can be understood in terms of the set-based multiples proposed by Alain Badiou:

@@

Patel observed that the 'Have you forgotten' lists matter to people, and we can see why in the light of this proliferation of things in the list. A `r total` list of items is quite hard to list let alone read. 
Patel said that some people log on and go straight to the checkout, allowing the recommender system to supply them with the suggestions that they then move to the shopping basket.
This is the online equivalent of using an old shopping list and only adding new items when needed. 
It is also a strategy to avoid having to view too many lists, since the only way Tesco can show people what they might buy is by showing them lists of things (e.g. chicken products), but these lists are often too long to work with. For instance, Tesco has more than 1000 chicken products on its site.   
People inhabit lists and work with them no matter how they are made; thus vindicating ethnomethodologists assertion that social order is the product of ongoing sense-making practices. 
But it also suggests a transformation in the way that lists exist in the world. They are less the product of a discrete agent than an effect of an environment teeming with possible combinations and complementaries, a surplus of possible purchases that the supermarket needs to continually maintain but also filter in order to make it possible for something to happen.

@@

Patel described the data they on grocery purchases. 'We have 52 weeks of purchase data for each customer.' 
We know that the construction of lists has been transformed by recommender systems.
These transformations have been widely discussed in the context of big data, where questions of who has how much data are usually most prominent.
In terms of the ways that lists order the world, the amount of data is not so amount as the way in which things in the world take on a list-like existence. 
I want to dwell on this idea for a while. We have seen two different ways in which lists might be constructed in a recommender system. 
The first uses history and existence of previous lists, including the customer's own list, to populate a new list. 
The second uses many shopping lists to identify associations between things and then uses the strongest associations to recommend additions to a current list. 
The 'relevancy model,' however, tries to do something different. 
The problem that the proliferation of things on the shelves of supermarket or grocery warehouse confronts is partly combinatorial.
With  `r total` products (actually Patel mentions 200,000 products) can be combined in truly astronomical numbers of combinations.
If a typical shopping list has 20 items, then there are `r format(choose(total, 20), 2)` possible combinations.
Some of this vast number of possible shopping lists is highly unlikely in terms of shopping lists and histories of shopping, but potentially highly interesting and value as recommendation.

@@

The fact that so little was said about the models themselves suggests that they are now largely taken for granted in recommender systems. The models she mentioned are all classifiers, predictive models that treat a set of variables in a datasets as a vector space in which good classification means finding a way of defining a decision boundary between points that belong to different classes of things. Here the classes of things are binary: `recommended` or `not recommended.` Some classifier models, including logistic regression, generate a probability rather than simply a classification.

@@

Predictive models, we might say, encounter their own surplus of sensibility in the data as they search for the most accurate predictions. A  very brief comment from Patel -- 'we have lots of zeros' -- points to the problem here. Each of the millions of Tesco online grocery shoppers has bought some selection of the 200,000 products. But the product-customer matrix, the basic vector-space in which all recommender systems operate, remain very sparse and unpopulated.

@@

The discussion of imperfection was brief - it incorporated the notion of recommendations weighted by price and the notion of precision (or the proportion of products actually purchased.). Patel simply said that this metric was imperfect, but without elaborating why. Instead she moved to describe some of the ways in which the recommender system worked in practice.

@@

Random allocation of customers to the four categories (Test A, Test B, Control A, Control B) adds another layer of probability to the recommender system. This is done in the name of statistical validation of the effects or the uplift of the model. It also suggests that the effects of the model cannot be known in advance -- they can only be observed experimentally. Does this experimentality of the list point more to the propensities that Hansen says saturate experience?

It is perhaps significant that this random allocation of customers to control and test groups occurs without any relation to the particular profile or propensities of the customer. It seeks to statistically validate the effects -- the *uplift* -- of the model on  conversion events by allowing the effects of the model on what people do to be measured. The lift or uplift refers to a change in events associated with the same groups of people. This effectively is an experiment in inhabiting two different worlds.  This sets up a feedback look between the predictive model, the logistic regression, and the operation of the recommendation system.[^122] Once the experimental uplift methodology is in place, the recommender system can constantly modify the recommendation list in light of subsequent data on the uplift. Predictions, then, are not bound by fixed parameters of the model.

@@

Hansen refers to an ontological transformation in which predictive media or recommender systems would be seen as the closed realisation of the actually open propensities of the world. By understanding the world in terms of propensities, he argues, we have a change to engage with contemporary predictive media.
